---
title: "Predicting .. from Spotify Music"
author: Alyssia Goodwin & Nolan Guzman & Arthur Diaz
date: "May 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<<<<<<< Updated upstream
#Data Input and Preprocessing 

=======
## Intro

We are using the "Now That's What I Call Music(U.S. releases)" data set from Kaggle to predict the danceability of a song. Danceability is described with how much a song will get people todance from the tempo, rhythm stability, beat, and regulatiry. The scale of danceability ranges from 0 to 1, 0 being the least danceable while 1 is the most danceable song. 



## Preprocessing 

Looking at the data there does not seem to be any null data within the database. From looking at the database it deo
>>>>>>> Stashed changes
```{r}
dat=read.csv("data.csv")

sum(is.null(dat))
dat$X = NULL
dat$loudness=abs(dat$loudness)
dat$song_title = NULL
```

#Data Exploration

```{r}
summary(dat)
```


<<<<<<< Updated upstream


=======
```{r}

plot(dat$tempo ~ dat$speechiness)
plot(dat)

plot(dat$energy~dat$loudness)
plot(dat$acousticness~dat$loudness)
plot(density(dat$energy))

```
>>>>>>> Stashed changes
